Supplementary Information to "Improving Models to Predict Holocellulose and Klason Lignin Contents for Peat Soil Organic Matter with Mid Infrared Spectra

### 2022-03-03 analysis/paper/006-paper-m-prediction-domain.Rmd
Update manuscript 1, update formatting fo references.

# binned spectra, best model
d_prediction_domain_hol_bin <-
  dplyr::bind_rows(
    # training data
    d_underfitting[[4]] %>% 
      dplyr::select(-kl, -hol) %>%
      tidyr::pivot_longer(cols = dplyr::everything(),
                          names_to = "variable_x", values_to = "value_x") %>%
      dplyr::mutate(
        wn = 
          d_underfitting_flat[[4]]$x %>%
          rep(nrow(d_underfitting[[4]])),
        sample_id = 
          seq_len(nrow(d_underfitting[[4]])) %>% 
          rep(each = nrow(d_underfitting_flat[[4]])) %>% 
          as.character(),
        data_type = "Training"
      ),
    # peat data
     d_peat_bin50 %>% 
      tidyr::pivot_longer(cols = dplyr::everything(),
                          names_to = "variable_x", values_to = "value_x") %>%
      dplyr::mutate(
        wn = 
          d_peat_bin50_flat$x %>%
          rep(nrow(d_peat_bin50)),
        sample_id = 
          seq_len(nrow(d_peat_bin50)) %>% 
          rep(each = nrow(d_peat_bin50_flat)) %>% 
          magrittr::add(nrow(d_underfitting[[4]])) %>%
          as.character(),
        data_type = "Peat"
      ),
    # vegetation data
     d_veg_bin50 %>% 
      tidyr::pivot_longer(cols = dplyr::everything(),
                          names_to = "variable_x", values_to = "value_x") %>%
      dplyr::mutate(
        wn = 
          d_veg_bin50_flat$x %>%
          rep(nrow(d_veg_bin50)),
        sample_id = 
          seq_len(nrow(d_veg_bin50)) %>% 
          rep(each = nrow(d_veg_bin50_flat)) %>% 
          magrittr::add(nrow(d_underfitting[[4]]) + nrow(d_peat_bin50)) %>%
          as.character(),
        data_type = "Vegetation"
      )
  ) %>%
  dplyr::group_by(data_type, wn) %>%
  dplyr::summarise(value_x_min = min(value_x),
                   value_x_max = max(value_x),
                   .groups = "keep") %>%
  dplyr::mutate(y_variable = "Holocellulose",
                model = "Binned")

@@

# binned spectra, best model
d_prediction_domain_kl_bin <-
  dplyr::bind_rows(
    # training data
    d_underfitting[[1]] %>% 
      dplyr::select(-kl, -hol) %>%
      tidyr::pivot_longer(cols = dplyr::everything(),
                          names_to = "variable_x", values_to = "value_x") %>%
      dplyr::mutate(
        wn = 
          d_underfitting_flat[[1]]$x %>%
          rep(nrow(d_underfitting[[1]])),
        sample_id = 
          seq_len(nrow(d_underfitting[[1]])) %>% 
          rep(each = nrow(d_underfitting_flat[[1]])) %>% 
          as.character(),
        data_type = "Training"
      ),
    # peat data
     d_peat_bin10 %>% 
      tidyr::pivot_longer(cols = dplyr::everything(),
                          names_to = "variable_x", values_to = "value_x") %>%
      dplyr::mutate(
        wn = 
          d_peat_bin10_flat$x %>%
          rep(nrow(d_peat_bin10)),
        sample_id = 
          seq_len(nrow(d_peat_bin10)) %>% 
          rep(each = nrow(d_peat_bin10_flat)) %>% 
          magrittr::add(nrow(d_underfitting[[1]])) %>%
          as.character(),
        data_type = "Peat"
      ),
    # vegetation data
     d_veg_bin10 %>% 
      tidyr::pivot_longer(cols = dplyr::everything(),
                          names_to = "variable_x", values_to = "value_x") %>%
      dplyr::mutate(
        wn = 
          d_veg_bin10_flat$x %>%
          rep(nrow(d_veg_bin10)),
        sample_id = 
          seq_len(nrow(d_veg_bin10)) %>% 
          rep(each = nrow(d_veg_bin10_flat)) %>% 
          magrittr::add(nrow(d_underfitting[[1]]) + nrow(d_peat_bin10)) %>%
          as.character(),
        data_type = "Vegetation"
      )
  ) %>%
  dplyr::group_by(data_type, wn) %>%
  dplyr::summarise(value_x_min = min(value_x),
                   value_x_max = max(value_x),
                   .groups = "keep") %>%
  dplyr::mutate(y_variable = "Klason lignin",
                model = "Binned")

@@

p3 <-
  d_prediction_domain_hol_bin %>%
  dplyr::mutate(
    data_type = factor(data_type, levels = data_type_levels)
  ) %>%
  ggplot(aes(x = wn, ymin = value_x_min, ymax = value_x_max, fill = data_type)) +
  geom_hline(yintercept = 0, color = "grey") +
  geom_ribbon(colour = NA, alpha = 1) +
  geom_vline(xintercept = 
               m_pars %>%
               dplyr::filter(stringr::str_detect(variable_y, "Holocellulose") & abs(median) >= 0.02) %>%
               dplyr::pull(wn),
             colour = "grey", linetype = 2) + 
  labs(x = expression("Bin wavenumber ["*cm^{-1}*"]"),
       y = "Predictor value",
       title = "Bin width = 50") +
  guides(fill = guide_legend(title = "Dataset")) +
  scale_fill_manual(
    breaks = data_type_levels, 
    values = colors_prediction_domain
  )

p4 <-
  d_prediction_domain_kl_bin %>%
  dplyr::mutate(
    data_type = factor(data_type, levels = data_type_levels)
  ) %>%
  ggplot(aes(x = wn, ymin = value_x_min, ymax = value_x_max, fill = data_type)) +
  geom_hline(yintercept = 0, color = "grey") +
  geom_ribbon(colour = NA, alpha = 1) +
  geom_vline(xintercept =  
               m_pars %>%
               dplyr::filter(stringr::str_detect(variable_y, "Klason lignin") & abs(median) >= 0.02) %>%
               dplyr::pull(wn),
             colour = "grey", linetype = 2) + # ---todo: remove hard coding of values
  labs(x = expression("Bin wavenumber ["*cm^{-1}*"]"),
       y = "Predictor value",
       title = "Bin width = 10") +
  guides(fill = guide_legend(title = "Dataset")) +
  scale_fill_manual(
    breaks = data_type_levels, 
    values = colors_prediction_domain
  )

# combine plots
pp1 <-
  (p1 | p2) +
  plot_annotation(title = "Original models")

pp2 <-
  p3 / p4 +
  plot_annotation(title = "Models using binned spectra")

p_prediction_domain <- 
  pp1 / pp2 +
  plot_annotation(tag_levels = "a", tag_prefix = "(", tag_suffix = ")") +
  plot_layout(heights = c(0.4, 2), guides = "collect") &
  theme(legend.position = "bottom",
        legend.box = "vertical")
```

>>>>

# binned spectra, best model
d_prediction_domain_hol_bin <-
  dplyr::bind_rows(
    # training data
    d_underfitting[[2]] %>% 
      dplyr::select(-kl, -hol) %>%
      tidyr::pivot_longer(cols = dplyr::everything(),
                          names_to = "variable_x", values_to = "value_x") %>%
      dplyr::mutate(
        wn = 
          d_underfitting_flat[[2]]$x %>%
          rep(nrow(d_underfitting[[2]])),
        sample_id = 
          seq_len(nrow(d_underfitting[[2]])) %>% 
          rep(each = nrow(d_underfitting_flat[[2]])) %>% 
          as.character(),
        data_type = "Training"
      ),
    # peat data
     d_peat_bin20 %>% 
      tidyr::pivot_longer(cols = dplyr::everything(),
                          names_to = "variable_x", values_to = "value_x") %>%
      dplyr::mutate(
        wn = 
          d_peat_bin20_flat$x %>%
          rep(nrow(d_peat_bin20)),
        sample_id = 
          seq_len(nrow(d_peat_bin20)) %>% 
          rep(each = nrow(d_peat_bin20_flat)) %>% 
          magrittr::add(nrow(d_underfitting[[2]])) %>%
          as.character(),
        data_type = "Peat"
      ),
    # vegetation data
     d_veg_bin20 %>% 
      tidyr::pivot_longer(cols = dplyr::everything(),
                          names_to = "variable_x", values_to = "value_x") %>%
      dplyr::mutate(
        wn = 
          d_veg_bin20_flat$x %>%
          rep(nrow(d_veg_bin20)),
        sample_id = 
          seq_len(nrow(d_veg_bin20)) %>% 
          rep(each = nrow(d_veg_bin20_flat)) %>% 
          magrittr::add(nrow(d_underfitting[[2]]) + nrow(d_peat_bin20)) %>%
          as.character(),
        data_type = "Vegetation"
      )
  ) %>%
  dplyr::group_by(data_type, wn) %>%
  dplyr::summarise(value_x_min = min(value_x),
                   value_x_max = max(value_x),
                   .groups = "keep") %>%
  dplyr::mutate(y_variable = "Holocellulose",
                model = "Binned")

@@

# binned spectra, best model
d_prediction_domain_kl_bin <-
  dplyr::bind_rows(
    # training data
    d_underfitting[[2]] %>% 
      dplyr::select(-kl, -hol) %>%
      tidyr::pivot_longer(cols = dplyr::everything(),
                          names_to = "variable_x", values_to = "value_x") %>%
      dplyr::mutate(
        wn = 
          d_underfitting_flat[[2]]$x %>%
          rep(nrow(d_underfitting[[2]])),
        sample_id = 
          seq_len(nrow(d_underfitting[[2]])) %>% 
          rep(each = nrow(d_underfitting_flat[[2]])) %>% 
          as.character(),
        data_type = "Training"
      ),
    # peat data
     d_peat_bin20 %>% 
      tidyr::pivot_longer(cols = dplyr::everything(),
                          names_to = "variable_x", values_to = "value_x") %>%
      dplyr::mutate(
        wn = 
          d_peat_bin20_flat$x %>%
          rep(nrow(d_peat_bin20)),
        sample_id = 
          seq_len(nrow(d_peat_bin20)) %>% 
          rep(each = nrow(d_peat_bin20_flat)) %>% 
          magrittr::add(nrow(d_underfitting[[2]])) %>%
          as.character(),
        data_type = "Peat"
      ),
    # vegetation data
     d_veg_bin20 %>% 
      tidyr::pivot_longer(cols = dplyr::everything(),
                          names_to = "variable_x", values_to = "value_x") %>%
      dplyr::mutate(
        wn = 
          d_veg_bin20_flat$x %>%
          rep(nrow(d_veg_bin20)),
        sample_id = 
          seq_len(nrow(d_veg_bin20)) %>% 
          rep(each = nrow(d_veg_bin20_flat)) %>% 
          magrittr::add(nrow(d_underfitting[[2]]) + nrow(d_peat_bin20)) %>%
          as.character(),
        data_type = "Vegetation"
      )
  ) %>%
  dplyr::group_by(data_type, wn) %>%
  dplyr::summarise(value_x_min = min(value_x),
                   value_x_max = max(value_x),
                   .groups = "keep") %>%
  dplyr::mutate(y_variable = "Klason lignin",
                model = "Binned")

@@

p3 <-
  d_prediction_domain_hol_bin %>%
  dplyr::mutate(
    data_type = factor(data_type, levels = data_type_levels)
  ) %>%
  ggplot(aes(x = wn, ymin = value_x_min, ymax = value_x_max, fill = data_type)) +
  geom_hline(yintercept = 0, color = "grey") +
  geom_ribbon(colour = NA, alpha = 1) +
  geom_vline(xintercept = 
               m_pars %>%
               dplyr::filter(stringr::str_detect(variable_y, "Holocellulose") & abs(median) >= 0.02) %>%
               dplyr::pull(wn),
             colour = "grey", linetype = 2) + 
  geom_vline(xintercept =  
               m_pars %>%
               dplyr::filter(stringr::str_detect(variable_y, "Klason lignin") & abs(median) >= 0.02) %>%
               dplyr::pull(wn),
             colour = "grey", linetype = 2) + 
  labs(x = expression("Bin wavenumber ["*cm^{-1}*"]"),
       y = "Predictor value",
       title = "Bin width = 20") +
  guides(fill = guide_legend(title = "Dataset")) +
  scale_fill_manual(
    breaks = data_type_levels, 
    values = colors_prediction_domain
  )

p4 <-
  d_prediction_domain_kl_bin %>%
  dplyr::mutate(
    data_type = factor(data_type, levels = data_type_levels)
  ) %>%
  ggplot(aes(x = wn, ymin = value_x_min, ymax = value_x_max, fill = data_type)) +
  geom_hline(yintercept = 0, color = "grey") +
  geom_ribbon(colour = NA, alpha = 1) +
  geom_vline(xintercept =  
               m_pars %>%
               dplyr::filter(stringr::str_detect(variable_y, "Klason lignin") & abs(median) >= 0.02) %>%
               dplyr::pull(wn),
             colour = "grey", linetype = 2) + 
  labs(x = expression("Bin wavenumber ["*cm^{-1}*"]"),
       y = "Predictor value",
       title = "Bin width = 10") +
  guides(fill = guide_legend(title = "Dataset")) +
  scale_fill_manual(
    breaks = data_type_levels, 
    values = colors_prediction_domain
  )

# combine plots
p_prediction_domain <-
  (p1 | p2 | p3) + 
  plot_annotation(tag_levels = "a", tag_prefix = "(", tag_suffix = ")") +
  plot_layout(widths = c(1, 1, 3), guides = "collect") &
  theme(legend.position = "bottom",
        legend.box = "vertical")

## old version wit two different plots for holocellulose and Klason lignin
## combine plots
#pp1 <-
#  (p1 | p2) +
#  plot_annotation(title = "Original models")
#
#pp2 <-
#  p3 / p4 +
#  plot_annotation(title = "Models using binned spectra")
#
#p_prediction_domain <- 
#  pp1 / pp2 +
#  plot_annotation(tag_levels = "a", tag_prefix = "(", tag_suffix = ")") +
#  plot_layout(heights = c(0.4, 2), guides = "collect") &
#  theme(legend.position = "bottom",
#        legend.box = "vertical")
```
